---
title: "CranioRNAseq- SVA and Voom Adjustment"
author: "Samantha Lapehn Young"
date: "5/05/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Male Craniosynostosis RNAseq: SVA and Voom Quality Weights Adjustment

This code performs the surrogate variable analysis and voom transformation for the single suture craniosynostosis RNAseq data for the male model. 

## Load Packages
```{r Packages}
library(tidyverse)
library(sva)
library(limma)
library(edgeR)
```


## Load Data
Load Data that was Filtered from Craniosynostis_MaleModel_Filtering.Rmd
```{r Load Data}
load("/Volumes/paquette_a/slapehn/CranioRNAseq/Male_CranioRNAseq/Data/Male_CranioRNAseq_FilteredNormalized.RData")
```

## Run SVA 
Creates surrogate variables that explain variability that is not explained by available covariates
```{r SVA}
#Step 1: Prepare model matrices (Full Model + Null Model)
mod <- model.matrix(~CranioCovariates$Phenotype +
                      CranioCovariates$Proband_Age +
                      CranioCovariates$Culture_Time +
                      CranioCovariates$Cohort + 
                      CranioCovariates$Origin)

#Null model (mod0) contains all covariates except Phenotype which is the covariate of interest
mod0 <- model.matrix(~1 + CranioCovariates$Proband_Age +
                      CranioCovariates$Culture_Time +
                      CranioCovariates$Cohort + 
                      CranioCovariates$Origin)

#Step 2: Run SVA
sv <- sva(cpm(y_Final, log = TRUE),mod, mod0)

#Step 3: Add surrogate variables (sv) to model
mod<-cbind(mod,sv$sv)
  dim(mod)
```

## Voom Transform Data
Voom transform the data to use in the limma differential gene expression pipeline
```{r Voom}
v<-voom(y_Final, mod,plot=TRUE)
save(v, CranioCovariates, mod, file="/Volumes/paquette_a/slapehn/CranioRNAseq/Male_CranioRNAseq/Data/Male_CranioRNAseq_VoomAdjustedData.RData")
```
